Rank-based Training-Free NAS Algorithm

### 2023-09-15 paper/manuscript.tex
update paper

\begin{abstract}
    The training-free neural architecture search (NAS) typically uses one  
    score function to evaluate how a neural network architecture performs. 
    However, every architecture has unique characteristics. To accurately 
    evaluate an architecture, the score function should take the different 
    aspects of an neural architecture into account. 
    To achieve the goal, we propose a rank-based training-free NAS algorithm, 
    which combines three ``complementary'' training-free score functions to 
	evaluate an architecture. 
	NAS without training score function measures the difference 
    of the activation patterns between a batch of data passing through an 
    architecture, i.e., the image distinction ability. 
	Noise immunity measures the difference between the pooling layers 
    of an architecture by passing an image and the same image with Gaussian noise, 
    i.e., the image generalization ability. 
	And LogSynFlow measures the importance of connections of an architecture, 
    i.e., the trainability. 
    Through the three measurements, an architecture can be evaluated from 
    different aspects. With an integrated score function, the proposed 
    method, \palg{} applies simulated annealing algorithm to obtain 
    better performance on searching high-accuracy architectures. 
    To evaluate the performance of \palg{}, this paper compared 
    it with several NAS algorithms, including weight sharing methods, 
    non-weight sharing methods, and several state-of-the-art training-free 
    score functions. The ultimate results show that \palg{} outperforms most 
    training-free score functions and shows the robustness between different 
    search spaces.

@@

    \begin{table*}[htb]
        %\normalsize
        %\centering
        \newcommand{\z}{\phantom{0}}
        \caption{\textsc{Comparison of rank-based NAS and all the other NAS algorithms in NATS-Bench-SSS.}}
          \vspace{-\baselineskip}
        \begin{tabular}{@{}lccccccc@{}}\toprule
        Method & Search (s) & \multicolumn{2}{c}{CIFAR-10} & \multicolumn{2}{c}{CIFAR-100} & \multicolumn{2}{c}{ImageNet-16-120} \\ \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
                            &          & validation       & test             & validation       & test              & validation       & test \\ \midrule
                            &          &                  &                  & \textbf{Non-weight sharing} &        &                  &                  \\
        AREA                & $12,000$ & $84.62 \pm 0.41$ & $93.16 \pm 0.16$ & $59.24 \pm 1.11$ & $69.56 \pm 0.96$  & $37.58 \pm 1.09$ & $45.30 \pm 0.91$ \\
        REA                 & $12,000$ & $90.37 \pm 0.20$ & $93.22 \pm 0.16$ & $70.23 \pm 0.50$ & $70.11 \pm 0.61$  & $45.30 \pm 0.69$ & $45.94 \pm 0.92$ \\
        RS                  & $12,000$ & $90.10 \pm 0.26$ & $93.03 \pm 0.25$ & $69.57 \pm 0.57$ & $69.72 \pm 0.61$  & $45.01 \pm 0.74$ & $45.42 \pm 0.86$ \\
        RL                  & $12,000$ & $90.25 \pm 0.23$ & $93.16 \pm 0.21$ & $69.84 \pm 0.59$ & $69.96 \pm 0.57$  & $45.06 \pm 0.77$ & $45.71 \pm 0.93$ \\
        BOHB                & $12,000$ & $90.07 \pm 0.28$ & $93.01 \pm 0.24$ & $69.75 \pm 0.60$ & $69.90 \pm 0.60$  & $45.11 \pm 0.69$ & $45.56 \pm 0.81$ \\ \midrule
                            &          &                  &                  & \textbf{Training-free} &             &                  &                  \\
        RS-NI (N=1,000)         & $293$& $89.55 \pm 0.16$ & $92.56 \pm 0.21$ & $67.67 \pm 0.91$ & $67.61 \pm 1.06$  & $43.19 \pm 0.73$ & $43.22 \pm 0.53$ \\ 
        RS-NASWOT (N=1,000)     & $208$& $89.25 \pm 0.44$ & $92.21 \pm 0.31$ & $65.09 \pm 2.89$ & $64.61 \pm 3.16$  & $41.29 \pm 2.32$ & $41.35 \pm 2.31$ \\ 
        RS-LogSynFlow (N=1,000) & $123$& $89.60 \pm 0.11$ & $92.61 \pm 0.18$ & $68.23 \pm 0.49$ & $68.30 \pm 0.54$  & $43.58 \pm 0.58$ & $43.48 \pm 0.40$ \\ 
        RS-rank (ours) (N=1000) & $612$& $89.59 \pm 0.14$ & $92.51 \pm 0.17$ & $67.56 \pm 0.75$ & $67.55 \pm 0.82$  & $43.36 \pm 0.62$ & $43.27 \pm 0.45$ \\ 
        \palg{} (ours)          & $725$& $90.33 \pm 0.15$ & $93.26 \pm 0.24$ & $69.89 \pm 0.48$ & $70.24 \pm 0.42$  & $45.41 \pm 0.41$ & $46.43 \pm 0.75$ \\ \midrule
        %SA-NI                  &      & $90.24 \pm 0.10$ & $93.20 \pm 0.19$ & $69.83 \pm 0.56$ & $70.10 \pm 0.40$  & $45.49 \pm 0.50$ & $45.98 \pm 0.89$
        %SA-NASWOT              &      & $90.21 \pm 0.30$ & $93.16 \pm 0.28$ & $69.16 \pm 1.28$ & $69.27 \pm 1.46$  & $44.89 \pm 1.27$ & $45.49 \pm 1.33$
        Optimal             &          & $90.71$          & $93.65$          & $70.92$          & $71.34$           & $46.73$          & $47.40$          \\ \bottomrule
        \end{tabular}
        \label{table:overall_sss}
          \vspace{-\baselineskip}
    \end{table*}

@@

        RS-NI (N=1,000)         & $592$  & $90.03 \pm \z{0.30}$ & $93.21 \pm 0.63$ & $70.27 \pm 0.70$ & $70.40 \pm 0.84$ & $42.92 \pm 2.18$ & $43.34 \pm 2.43$ \\ 
        RS-NASWOT (N=1,000)     & $201$  & $89.63 \pm \z{0.94}$ & $92.80 \pm 1.07$ & $69.76 \pm 1.25$ & $69.66 \pm 1.23$ & $43.51 \pm 2.90$ & $43.84 \pm 2.93$ \\ 
        RS-LogSynFlow (N=1,000) & $90$   & $89.67 \pm \z{0.89}$ & $92.92 \pm 0.70$ & $69.19 \pm 1.86$ & $68.98 \pm 1.97$ & $39.93 \pm 6.34$ & $40.23 \pm 6.45$ \\ 
        RS-rank (ours) (N=1000) & $887$  & $90.07 \pm \z{0.41}$ & $93.32 \pm 0.42$ & $70.06 \pm 1.34$ & $70.10 \pm 1.45$ & $43.68 \pm 2.11$ & $44.07 \pm 2.19$ \\ 

@@

@@

        %SA-NI                  &        & $90.06 \pm \z{0.29}$ & $93.38 \pm 0.50$ & $70.41 \pm 0.74$ & $70.34 \pm 1.34$ & $42.69 \pm 1.87$ & $43.22 \pm 1.93$
        %SA-NASWOT              &        & $89.11 \pm \z{1.22}$ & $92.22 \pm 1.32$ & $69.38 \pm 1.23$ & $69.48 \pm 1.46$ & $41.77 \pm 3.96$ & 

@@

    % tss
    In NATS-Bench-TSS \cite{Dong_2021}, \xtab{table:overall_201} compares \palg{} with the other 
    non-weight sharing algorithms, six weight sharing algorithms, three different training-free 
    score functions with RS, i.e., RS-NI, RS-NASWOT, and RS-LogSynFlow and one 
    training-free method, TE-NAS. Same as the experiment of NATS-Bench-SSS, the searching 
    architectures are identical for the RS with score functions. Non-weight sharing algorithms 
    like REA and AREA outperform all the other search algorithms, though the time cost is 
    tremendous. Among the four training-free score functions with random search, RS-rank has the 
    highest accuracy in CIFAR10 and ImageNet16. But in CIFAR100, RS-NI has the highest accuracy. 
    In CIFAR10 validation dataset, RS-NI is the most stable, RS-rank the second, RS-LogSynFlow the 
    third and RS-NASWOT the last. In CIFAR10 test dataset, RS-rank is the most stable, then is RS-NI, 
    RS-LogSynFlow, RS-NASWOT. In CIFAR100, RS-NI is the most stable, RS-NASWOT the second but 
    has lower accuracy, RS-rank the third, the last is RS-LogSynFlow. In ImageNet16, RS-rank is the 
    most stable, then is RS-NI, RS-NASWOT, RS-LogSynFlow. According to the results, RS-rank does 
    integrate the score functions and use them to evaluate an architecture by different characteristics. 
    With the help of search algorithm, \palg{} outperforms all the others training-free methods, 
    and obtains a better stability. In CIFAR10 and CIFAR100, the standard deviation of \palg{} is lower 
    than all the non-weight sharing algorithms. And in ImageNet16, the standard deviation of \palg{} is 
    lower than all the training-free methods.

@@

    \begin{table}[t]
        %\normalsize
        %\centering
        \newcommand{\z}{\phantom{0}}
        \caption{\textsc{Comparison of rank-based NAS and all the other NAS algorithms in NAS-Bench-101.}}
        \vspace{-\baselineskip}
        \resizebox{0.5 \textwidth}{!}{
        \begin{tabular}{@{}lrcc@{}}\toprule
            Method              & Search (s) & \multicolumn{2}{c}{CIFAR-10}        \\ \cmidrule(lr){3-4}
                                &            & validation       & test             \\ \midrule
                                &            & \textbf{Non-weight sharing} &       \\
            AREA                & $12,000$   & $93.67 \pm 0.48$ & $93.02 \pm 0.48$ \\
            REA                 & $12,000$   & $93.63 \pm 0.50$ & $93.02 \pm 0.52$ \\ \midrule
                                &            & \textbf{Weight sharing}     &       \\
            DARTS-V1            & $20,483$   & $83.50 \pm 0.11$ & $83.74 \pm 0.03$ \\
            ENAS                & $22,325$   & $93.83 \pm 0.28$ & $93.34 \pm 0.26$ \\ \midrule
                                &            & \textbf{Training-free}      &       \\
            RS-NI (N=1,000)         & $476$  & $92.89 \pm 2.10$ & $92.55 \pm 1.33$ \\ 
            RS-NASWOT (N=1,000)     & $127$  & $92.27 \pm 6.60$ & $92.10 \pm 4.44$ \\ 
            RS-LogSynFlow (N=1,000) & $76$   & $91.81 \pm 1.52$ & $91.21 \pm 1.71$ \\ 
            RS-rank (ours) (N=1000) & $679$  & $92.44 \pm 1.42$ & $91.94 \pm 2.01$ \\
            RS-NI (N=5,000)         & $2390$ & $92.84 \pm 1.55$ & $92.35 \pm 2.69$ \\ 
            RS-NASWOT (N=5,000)     & $635$  & $91.67 \pm 4.99$ & $90.60 \pm 6.64$ \\ 
            RS-LogSynFlow (N=5,000) & $408$  & $91.66 \pm 1.72$ & $91.17 \pm 1.86$ \\ 
            RS-rank (ours) (N=5,000)& $3434$  & $92.87 \pm 0.97$ & $92.32 \pm 1.47$ \\ 
            \palg{} (ours)          & $595$  & $93.16 \pm 0.97$ & $92.77 \pm 1.03$ \\ \midrule
            %SA-NI                  &        & $92.38 \pm 3.50$ & $92.71 \pm 1.97$
            %SA-NASWOT
            Optimal             &            & $95.12$          & $94.32$          \\ \bottomrule
            \end{tabular}
        }
        \label{table:overall_101}
        \vspace{-\baselineskip}
    \end{table}

>>>>

\begin{abstract}
    The training-free neural architecture search (NAS) typically uses one  
    score function to evaluate how a neural network architecture performs. 
    However, every architecture has unique characteristics. To accurately 
    evaluate an architecture, the score function should take the different 
    aspects of a neural architecture into account. 
    To achieve the goal, we propose a rank-based training-free NAS algorithm, 
    which combines three ``complementary'' training-free score functions to 
	evaluate an architecture. 
	NAS without training score function measures the difference 
    of the activation patterns between a batch of data passing through an 
    architecture, i.e., the image distinction ability. 
	Noise immunity measures the difference between the pooling layers 
    of an architecture by passing an image and the same image with Gaussian noise, 
    i.e., the image generalization ability. 
	And LogSynFlow measures the importance of connections of an architecture, 
    i.e., the trainability. 
    Through the three measurements, an architecture can be evaluated from 
    different aspects. With an integrated score function, the proposed 
    method, \palg{} applies simulated annealing algorithm to obtain 
    better performance on searching high-accuracy architectures. 
    To evaluate the performance of \palg{}, this paper compared 
    it with several NAS algorithms, including weight sharing methods, 
    non-weight sharing methods, and several state-of-the-art training-free 
    score functions. The ultimate results show that \palg{} outperforms most 
    training-free score functions and shows the robustness between different 
    search spaces.

@@

    \begin{table*}[htb]
        %\normalsize
        %\centering
        \newcommand{\z}{\phantom{0}}
        \caption{\textsc{Comparison of rank-based NAS and all the other NAS algorithms in NATS-Bench-SSS.}}
          \vspace{-\baselineskip}
        \begin{tabular}{@{}lccccccc@{}}\toprule
        Method & Search (s) & \multicolumn{2}{c}{CIFAR-10} & \multicolumn{2}{c}{CIFAR-100} & \multicolumn{2}{c}{ImageNet-16-120} \\ \cmidrule(lr){3-4}\cmidrule(lr){5-6}\cmidrule(lr){7-8}
                            &          & validation       & test             & validation       & test              & validation       & test \\ \midrule
                            &          &                  &                  & \textbf{Non-weight sharing} &        &                  &                  \\
        AREA                & $12,000$ & $84.62 \pm 0.41$ & $93.16 \pm 0.16$ & $59.24 \pm 1.11$ & $69.56 \pm 0.96$  & $37.58 \pm 1.09$ & $45.30 \pm 0.91$ \\
        REA                 & $12,000$ & $90.37 \pm 0.20$ & $93.22 \pm 0.16$ & $70.23 \pm 0.50$ & $70.11 \pm 0.61$  & $45.30 \pm 0.69$ & $45.94 \pm 0.92$ \\
        RS                  & $12,000$ & $90.10 \pm 0.26$ & $93.03 \pm 0.25$ & $69.57 \pm 0.57$ & $69.72 \pm 0.61$  & $45.01 \pm 0.74$ & $45.42 \pm 0.86$ \\
        RL                  & $12,000$ & $90.25 \pm 0.23$ & $93.16 \pm 0.21$ & $69.84 \pm 0.59$ & $69.96 \pm 0.57$  & $45.06 \pm 0.77$ & $45.71 \pm 0.93$ \\
        BOHB                & $12,000$ & $90.07 \pm 0.28$ & $93.01 \pm 0.24$ & $69.75 \pm 0.60$ & $69.90 \pm 0.60$  & $45.11 \pm 0.69$ & $45.56 \pm 0.81$ \\ \midrule
                            &          &                  &                  & \textbf{Training-free} &             &                  &                  \\
        RS-NI (N=1,000)         & $265$& $89.55 \pm 0.16$ & $92.56 \pm 0.21$ & $67.67 \pm 0.91$ & $67.61 \pm 1.06$  & $43.19 \pm 0.73$ & $43.22 \pm 0.53$ \\ 
        RS-NASWOT (N=1,000)     & $187$& $89.25 \pm 0.44$ & $92.21 \pm 0.31$ & $65.09 \pm 2.89$ & $64.61 \pm 3.16$  & $41.29 \pm 2.32$ & $41.35 \pm 2.31$ \\ 
        RS-LogSynFlow (N=1,000) & $127$& $89.60 \pm 0.11$ & $92.61 \pm 0.18$ & $68.23 \pm 0.49$ & $68.30 \pm 0.54$  & $43.58 \pm 0.58$ & $43.48 \pm 0.40$ \\ 
        RS-rank (ours) (N=1000) & $567$& $89.59 \pm 0.14$ & $92.51 \pm 0.17$ & $67.56 \pm 0.75$ & $67.55 \pm 0.82$  & $43.36 \pm 0.62$ & $43.27 \pm 0.45$ \\ 
        %SA-NI                   & $214$& $90.24 \pm 0.10$ & $93.20 \pm 0.19$ & $69.83 \pm 0.56$ & $70.10 \pm 0.40$  & $45.49 \pm 0.50$ & $45.98 \pm 0.89$ \\
        %SA-NASWOT               & $242$& $90.21 \pm 0.30$ & $93.16 \pm 0.28$ & $69.16 \pm 1.28$ & $69.27 \pm 1.46$  & $44.89 \pm 1.27$ & $45.49 \pm 1.33$ \\
        %SA-LogSynFlow           & $176$& $90.71 \pm 0.00$ & $93.40 \pm 0.00$ & $70.30 \pm 0.04$ & $70.72 \pm 0.00$  & $44.73 \pm 0.00$ & $47.17 \pm 0.00$ \\
        \palg{} (ours)          & $725$& $90.33 \pm 0.15$ & $93.26 \pm 0.24$ & $69.89 \pm 0.48$ & $70.24 \pm 0.42$  & $45.41 \pm 0.41$ & $46.43 \pm 0.75$ \\ \midrule
        Optimal             &          & $90.71$          & $93.65$          & $70.92$          & $71.34$           & $46.73$          & $47.40$          \\ \bottomrule
        \end{tabular}
        \label{table:overall_sss}
          \vspace{-\baselineskip}
    \end{table*}

@@

        RS-NI (N=1,000)         & $352$  & $90.03 \pm \z{0.30}$ & $93.21 \pm 0.63$ & $70.27 \pm 0.70$ & $70.40 \pm 0.84$ & $42.92 \pm 2.18$ & $43.34 \pm 2.43$ \\ 
        RS-NASWOT (N=1,000)     & $241$  & $89.63 \pm \z{0.94}$ & $92.80 \pm 1.07$ & $69.76 \pm 1.25$ & $69.66 \pm 1.23$ & $43.51 \pm 2.90$ & $43.84 \pm 2.93$ \\ 
        RS-LogSynFlow (N=1,000) & $155$  & $89.67 \pm \z{0.89}$ & $92.92 \pm 0.70$ & $69.19 \pm 1.86$ & $68.98 \pm 1.97$ & $39.93 \pm 6.34$ & $40.23 \pm 6.45$ \\ 
        RS-rank (ours) (N=1000) & $747$  & $90.07 \pm \z{0.41}$ & $93.32 \pm 0.42$ & $70.06 \pm 1.34$ & $70.10 \pm 1.45$ & $43.68 \pm 2.11$ & $44.07 \pm 2.19$ \\ 

@@

        %SA-NI                   & $319$  & $90.06 \pm \z{0.29}$ & $93.38 \pm 0.50$ & $70.41 \pm 0.74$ & $70.34 \pm 1.34$ & $42.69 \pm 1.87$ & $43.22 \pm 1.93$ \\
        %SA-NASWOT               & $273$  & $89.11 \pm \z{1.22}$ & $92.22 \pm 1.32$ & $69.38 \pm 1.23$ & $69.48 \pm 1.46$ & $41.77 \pm 3.96$ & $42.13 \pm 4.11$ \\
        %SA-LogSynFlow           & $213$  & $89.57 \pm \z{1.14}$ & $93.00 \pm 0.81$ & $69.30 \pm 1.86$ & $69.57 \pm 1.41$ & $40.42 \pm 4.61$ & $40.79 \pm 4.77$ \\

@@

@@

    % tss
    In NATS-Bench-TSS \cite{Dong_2021}, \xtab{table:overall_201} compares \palg{} with the other 
    non-weight sharing algorithms, six weight sharing algorithms, three different training-free 
    score functions with RS, i.e., RS-NI, RS-NASWOT, and RS-LogSynFlow and one 
    training-free method, TE-NAS. Same as the experiment of NATS-Bench-SSS, the searching 
    architectures are identical for the RS with score functions. Non-weight sharing algorithms 
    like REA and AREA outperform all the other search algorithms, though the time cost is 
    tremendous. Among the four training-free score functions with random search, RS-rank has the 
    highest accuracy in CIFAR10 and ImageNet16. But in CIFAR100, RS-NI has the highest accuracy. 
    In CIFAR10 validation dataset, RS-NI is the most stable, RS-rank the second, RS-LogSynFlow the 
    third and RS-NASWOT the last. In CIFAR10 test dataset, RS-rank is the most stable, then is RS-NI, 
    RS-LogSynFlow, RS-NASWOT. In CIFAR100, RS-NI is the most stable, RS-NASWOT the second but 
    has lower accuracy, RS-rank the third, the last is RS-LogSynFlow. In ImageNet16, RS-rank is the 
    most stable, then is RS-NI, RS-NASWOT, RS-LogSynFlow. According to the results, RS-rank does 
    integrate the score functions and use them to evaluate an architecture by different characteristics. 
    With the help of search algorithm, \palg{} outperforms all the others training-free methods, 
    and obtains a better stability. In CIFAR10 and CIFAR100, the standard deviation of \palg{} is lower 
    than all the non-weight sharing algorithms. And in ImageNet16, the standard deviation of \palg{} is 
    lower than all the training-free methods.

    \begin{table}[t]
        %\normalsize
        %\centering
        \newcommand{\z}{\phantom{0}}
        \caption{\textsc{Comparison of rank-based NAS and all the other NAS algorithms in NAS-Bench-101.}}
        \vspace{-\baselineskip}
        \resizebox{0.5 \textwidth}{!}{
        \begin{tabular}{@{}lrcc@{}}\toprule
            Method              & Search (s) & \multicolumn{2}{c}{CIFAR-10}        \\ \cmidrule(lr){3-4}
                                &            & validation       & test             \\ \midrule
                                &            & \textbf{Non-weight sharing} &       \\
            AREA                & $12,000$   & $93.67 \pm 0.48$ & $93.02 \pm 0.48$ \\
            REA                 & $12,000$   & $93.63 \pm 0.50$ & $93.02 \pm 0.52$ \\ \midrule
                                &            & \textbf{Weight sharing}     &       \\
            DARTS-V1            & $20,483$   & $83.50 \pm 0.11$ & $83.74 \pm 0.03$ \\
            ENAS                & $22,325$   & $93.83 \pm 0.28$ & $93.34 \pm 0.26$ \\ \midrule
                                &            & \textbf{Training-free}      &       \\
            RS-NI (N=1,000)         & $256$  & $92.84 \pm 2.43$ & $92.74 \pm 0.90$ \\ 
            RS-NASWOT (N=1,000)     & $137$  & $92.76 \pm 2.85$ & $91.84 \pm 3.68$ \\ 
            RS-LogSynFlow (N=1,000) & $93$   & $91.62 \pm 1.68$ & $91.42 \pm 1.82$ \\ 
            RS-rank (ours) (N=1000) & $485$  & $92.65 \pm 1.44$ & $92.25 \pm 1.31$ \\

            RS-NI (N=5,000)         & $2390$ & $92.84 \pm 1.55$ & $92.35 \pm 2.69$ \\ 
            RS-NASWOT (N=5,000)     & $635$  & $91.67 \pm 4.99$ & $90.60 \pm 6.64$ \\ 
            RS-LogSynFlow (N=5,000) & $408$  & $91.66 \pm 1.72$ & $91.17 \pm 1.86$ \\ 
            RS-rank (ours) (N=5,000)& $3434$ & $92.87 \pm 0.97$ & $92.32 \pm 1.47$ \\ 
            %SA-NI                   & $244$  & $92.38 \pm 3.50$ & $92.71 \pm 1.97$ \\
            %SA-NASWOT               & $136$  & $92.72 \pm 4.35$ & $91.95 \pm 3.62$ \\
            %SA-LogSynFlow           & $102$  & $92.21 \pm 1.50$ & $91.48 \pm 1.67$ \\
            \palg{} (ours)          & $595$  & $93.16 \pm 0.97$ & $92.77 \pm 1.03$ \\ \midrule

            Optimal             &            & $95.12$          & $94.32$          \\ \bottomrule
            \end{tabular}
        }
        \label{table:overall_101}
        \vspace{-\baselineskip}
    \end{table}

@@
